---
title: 'TSA: Forecasting Competition'
author: "Hannah Schofield & Sebrina Zu"
output:
  pdf_document: default
  html_document:
    df_print: paged
always_allow_html: yes
editor_options:
  chunk_output_type: console
  markdown:
    wrap: sentence
---

```{r package, message=FALSE, warning=FALSE, include = FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
#library(xlsx)
library(readxl)
```

## Data Wrangling

We began this project by loading in three datasets: hourly load, hourly temperature and hourly humidity.
Each dataset was manipulated to create figures for daily averages to be used in the forecasting process.
These datasets were then converted to time series objects, which was then subsetted into training and testing sets to inform the model.

```{r load, message=FALSE, include = FALSE}
#Loading data
load <- read_excel("/Users/hannahschofield/SchofieldZu_ENV790_TSA_Competition_S2023/Competition/Data/load.xlsx")
relative_humidity <- read_excel("/Users/hannahschofield/SchofieldZu_ENV790_TSA_Competition_S2023/Competition/Data/relative_humidity.xlsx")
temperature<-read_excel("/Users/hannahschofield/SchofieldZu_ENV790_TSA_Competition_S2023/Competition/Data/temperature.xlsx")
```

```{r data, message=FALSE, include = FALSE}
#convert hourly to daily averages 
load_daily <- load %>% 
  mutate(daily_average=rowMeans(load[,3:26],na.rm=TRUE))  %>% 
  select(date, daily_average)

#Temperature daily average
temperature_daily <- temperature %>% 
  mutate(average=rowMeans(temperature[,3:30],na.rm=TRUE))  %>% 
  select(date, average)%>%
  group_by(date) %>%
  summarize(daily_temp = sum(average))

#Humidity daily average
humidity_daily <- relative_humidity %>% 
  mutate(average_h=rowMeans(relative_humidity[,3:30],na.rm=TRUE))  %>% 
  select(date, average_h)%>%
  group_by(date) %>%
  summarize(daily_humid = sum(average_h))

```

### ACF and PACF

We first visualized the data, its residuals, and the pattern in the time series object.

```{r acfs, message = FALSE}
#ACF and PACF Plot
par(mfrow=c(1,2))
ACF_Plot <- Acf(load_daily$daily_average, lag = 40, plot = TRUE,main="")
PACF_Plot <- Pacf(load_daily$daily_average, lag = 40, plot = TRUE,main="")
par(mfrow=c(1,1))

```

```{r ts, include = FALSE}
#convert to time series 
ts_avgDaily <- msts(load_daily$daily_average, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,1,1))
autoplot(ts_avgDaily)
```

```{r testandtrain, echo = FALSE}
#create test and train data subsets for forecasting
n_for = 365
ts_avgDaily_train <- subset(ts_avgDaily,end = length(ts_avgDaily)-n_for)

ts_avgDaily_test <- subset(ts_avgDaily,start = length(ts_avgDaily)-n_for)

autoplot(ts_avgDaily_train)
autoplot(ts_avgDaily_test)

```

## Exogenous Data

```{r exog, include = FALSE}
df_list = list(temperature_daily,humidity_daily,load_daily) 
exog_data<-df_list %>% reduce(full_join, by='date')
exog_ts<-msts(exog_data[,2:3], 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,1,1))
n_for = 365
exog_train <- head(exog_ts, n=nrow(exog_ts)-n_for)

exog_test <- tail(exog_ts,n=n_for+1)

```

## Forecast on Training Data

### STL + ETS

```{r STL+ETS, echo = FALSE}
ETS_fit <-  stlf(ts_avgDaily_train,h=365)

#Plot foresting results
#autoplot(ETS_fit) + ylab("Load")

#Plot model + observed data
#autoplot(ts_avgDaily) +
  #autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  #ylab("Load")

ETS_fit_2010<-stlf(ts_avgDaily,h=59)

autoplot(ETS_fit_2010, ylab = "Load", main = "ETS 2010 vs Observed Data")
```

### ARIMA + Fourier

```{r ARIMA+Fourier, echo = FALSE}

ARIMA_Four_fit <- auto.arima(ts_avgDaily_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_avgDaily_train, 
                                          K=c(2,12)))

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(ts_avgDaily_train,
                                        K=c(2,12),
                                        h=365),
                                        h=365) 

#Plot foresting results
autoplot(ARIMA_Four_for, ylab = "Load", main = "ARIMA + Fourier Model")

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "ARIMA + Fourier vs Observed Data") +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) 

```

### Neural Network

```{r Neural Network, echo = FALSE}
NN_fit <- nnetar(ts_avgDaily_train,p=1,P=0,xreg=fourier(ts_avgDaily_train, K=c(2,12)))

NN_for <- forecast(NN_fit, h=365,xreg=fourier(ts_avgDaily_train, K=c(2,12),h=365))

#Plot foresting results
autoplot(NN_for, ylab = "Load", main = "Neural Network Forecast")

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "Neural Network vs Observed Data") +
  autolayer(NN_for, series="Neural Network",PI=FALSE)

```

### TBATS

```{r TBATS, echo = FALSE}
TBATS_fit <- tbats(ts_avgDaily_train)

TBATS_for <- forecast(TBATS_fit, h=n_for)

#Plot foresting results
autoplot(TBATS_for, ylab = "Load", main = "TBATS Model Forecast") 

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "TBATS vs Observed Data") +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)

```

### SARIMA

```{r SARIMA, echo = FALSE}
SARIMA_autofit <- auto.arima(ts_avgDaily_train)

SARIMA_for <- forecast(SARIMA_autofit,h=n_for)

#Plot foresting results
autoplot(SARIMA_for, ylab = "Load", main = "SARIMA Forecast") 

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "SARIMA vs Observed Data") +
  autolayer(SARIMA_for, series="SARIMA",PI=FALSE)
```

### ARIMA with Exogenous Variables

```{r ARIMAX with Exogenous Variables, echo = FALSE, warning = FALSE}
ARIMAX_fit_train <- auto.arima(ts_avgDaily_train, 
                             seasonal=FALSE, 
                             lambda=0,
                            xreg=cbind(exog_train[,1],exog_train[,2],fourier(ts_avgDaily_train,K=c(2,6))))

Exog_fit_temp_train<-auto.arima(exog_train[,1])
Exog_fit_humid_train<-auto.arima(exog_train[,2])
Temp_for_train<-forecast(Exog_fit_temp_train,h=365)
Humid_for_train<-forecast(Exog_fit_humid_train,h=365)


ARIMAX_Four_for_train <- forecast(ARIMAX_fit_train,                      xreg=cbind(Temp_for_train$mean,Humid_for_train$mean,fourier(ts_avgDaily_train,K=c(2,6),h=365)), h=365) 
#Plot foresting results
autoplot(ARIMAX_Four_for_train, ylab = "Load", main = "ARIMAX Forecast") 

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "ARIMAX vs Observed Data") +
  autolayer(ARIMAX_Four_for_train, series="ARIMAX",PI=FALSE)
```

### Neural Network with Exogenous Variables

```{r Neural Network + Exogenous Variables, echo = FALSE, warning = FALSE}
NNX_fit_train <- nnetar(ts_avgDaily_train,p=1,P=0,xreg=cbind(exog_train[,1],exog_train[,2],fourier(ts_avgDaily_train,K=c(2,6))))

NN_fit_temp_train<-nnetar(exog_train[,1])
NN_fit_humid_train<-nnetar(exog_train[,2])
NN_Temp_for_train<-forecast(NN_fit_temp_train,h=365)
NN_Humid_for_train<-forecast(NN_fit_humid_train,h=365)

NNX_for_train <- forecast(NNX_fit_train,xreg=cbind(NN_Temp_for_train$mean,NN_Humid_for_train$mean,fourier(ts_avgDaily_train,K=c(2,6),h=365)),h=365)

#Plot foresting results
autoplot(NNX_for_train, ylab = "Load", main = "Neural Network Exogenous Forecast") 

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "Neural Network Exogenous vs Observed Data") +
  autolayer(NNX_for_train, series="ARIMAX",PI=FALSE)
```

## Comparing Model Scores and Forecasts

```{r scores, echo = FALSE}
#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean,ts_avgDaily_test)  

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_Four_for$mean,ts_avgDaily_test)

# Model 3:  Neural Network 
NN_scores <- accuracy(NN_for$mean,ts_avgDaily_test)

## Model 4:  TBATS 
TBAT_scores <- accuracy(TBATS_for$mean,ts_avgDaily_test)

# Model 5:  SARIMA 
SARIMA_scores <- accuracy(SARIMA_for$mean,ts_avgDaily_test)

# Model 6:  SSES
#SSES_scores <- accuracy(SSES_seas$forecast,ts_avgDaily_test)

# Model 7:  BSM 
#SS_scores <- accuracy(SS_for$mean,ts_avgDaily_test)

# Model 8:  ARIMAX
ARIMAX_scores<- accuracy(ARIMAX_Four_for_train$mean,ts_avgDaily_test)

#Model 9: NNX
NNX_scores<- accuracy(NNX_for_train$mean,ts_avgDaily_test)
```

```{r MAPE, echo = FALSE}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, NN_scores,TBAT_scores,SARIMA_scores,ARIMAX_scores,NNX_scores))

row.names(scores) <- c("STL+ETS", "ARIMA+Fourier", "NN","TBAT","SARIMA","ARIMAX","NNX")

#choose model with lowest MAPE
best_model_index <- which.min(scores[,"MAPE"])
cat("The best model by MAPE is:", row.names(scores[best_model_index,]))  

```

```{r compare, echo = FALSE}
autoplot(ts_avgDaily_test, ylab = "Load", main = "Model Comparisons") +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(NN_for,PI=FALSE, series="NN") +
  guides(colour=guide_legend(title="Forecast"))

```

## Forecasting on 2010 Data (Final Forecast Result)

### Neural Network Forecast

```{r Neural Network2, echo = FALSE}
NN_fit_2010 <- nnetar(ts_avgDaily,p=1,P=0,xreg=fourier(ts_avgDaily, K=c(2,12)))

NN_for_2010 <- forecast(NN_fit_2010, h=59,xreg=fourier(ts_avgDaily, K=c(2,12),h=59))

autoplot(ts_avgDaily, ylab = "Load",main = "Neural Network vs Observed Data") +
  autolayer(NN_for_2010, series="NN",PI=FALSE) 

```

### ARIMA + Fourier Forecast

```{r ARIMA + Fourier2, echo = FALSE}
ARIMA_Four_fit_2010 <- auto.arima(ts_avgDaily, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_avgDaily, 
                                          K=c(2,12)))

ARIMA_Four_for_2010 <- forecast(ARIMA_Four_fit_2010,
                           xreg=fourier(ts_avgDaily,
                                        K=c(2,12),
                                        h=59),
                                        h=59) 

#Plot foresting results
autoplot(ARIMA_Four_for, ylab="Load", main = "ARIMA Fourier Forecast")

#Plot model + observed data
autoplot(ts_avgDaily, ylab = "Load", main = "ARIMA+Fourier vs Observed Data") +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE)

```

### ARIMA with Exogenous Variable Forecast

```{r ARIMAX (with Exogenous Variables)2, echo = FALSE, warning = FALSE}
ARIMAX_fit <- auto.arima(ts_avgDaily, 
                             seasonal=FALSE, 
                             lambda=0,        xreg=cbind(exog_ts[,1],exog_ts[,2],fourier(ts_avgDaily,K=c(2,12)))
                         )

Exog_fit_temp<-auto.arima(exog_ts[,1])
Exog_fit_humid<-auto.arima(exog_ts[,2])
Temp_for<-forecast(Exog_fit_temp,h=59)
Humid_for<-forecast(Exog_fit_humid,h=59)

ARIMAX_Four_for <- forecast(ARIMAX_fit,                      xreg=cbind(Temp_for$mean,Humid_for$mean,fourier(ts_avgDaily,K=c(2,12),h=59)),h=59) 

#Plot foresting results
autoplot(ARIMAX_Four_for, ylab="Load", main = "ARIMAX+Fourier Forecast")

autoplot(ts_avgDaily,  ylab="Load", main="ARIMAX+Fourier vs Observed Data")+
  autolayer(ARIMAX_Four_for, series="ARIMA_FOURIER",PI=FALSE) 
 

```

### Neural Network with Exogenous Variables Forecast

```{r Neural Network + Exogenous Variables2, echo = FALSE, warning = FALSE}
NNX_fit <- nnetar(ts_avgDaily,p=1,P=0,xreg=cbind(exog_ts[,1],exog_ts[,2],fourier(ts_avgDaily,K=c(2,6))))

NN_fit_temp<-nnetar(exog_ts[,1])
NN_fit_humid<-nnetar(exog_ts[,2])
NN_Temp_for<-forecast(NN_fit_temp,h=59)
NN_Humid_for<-forecast(NN_fit_humid,h=59)

NNX_for <- forecast(NNX_fit,xreg=cbind(NN_Temp_for$mean,NN_Humid_for$mean,fourier(ts_avgDaily,K=c(2,6),h=59)),h=59)

#Plot foresting results
autoplot(NNX_for, ylab="Load", main = "NNX Fourier Forecast")

autoplot(ts_avgDaily, ylab = "Load", main = "NNX vs Observed Data") +
  autolayer(NNX_for, series="NNX",PI=FALSE) 

```

### Submission Conversion

```{r submission1, results = FALSE, echo = FALSE}
#convert to submission format 

forecastRange <- seq(as.Date("2011-01-01"), as.Date("2011-02-28"), by="days")
forecastDat<- ARIMAX_Four_for$mean
submission <- data.frame(forecastRange, forecastDat)
colnames(submission)<-c("date","load")
write.csv(submission, file = "submission_schofield_zu_ARIMAX.csv",row.names = FALSE)

```

```{r submission2, results = FALSE, echo = FALSE}
forecastDat3<- NN_for_2010$mean
submission <- data.frame(forecastRange, forecastDat3)
colnames(submission)<-c("date","load")
write.csv(submission, file = "submission_schofield_zu_NN.csv",row.names = FALSE)

```

```{r submission3, results = FALSE, echo = FALSE}
#convert to submission format 

forecastRange <- seq(as.Date("2011-01-01"), as.Date("2011-02-28"), by="days")
forecastDat2<- NNX_for$mean
submission <- data.frame(forecastRange, forecastDat2)
colnames(submission)<-c("date","load")
write.csv(submission, file = "submission_schofield_zu_NNX.csv",row.names = FALSE)

```
